Blue Hunters: Bluetooth RSSI Locator Robots

### 2018-02-02 article/article.tex
Local refs budget
clean up all references

The HM-10 bluetooth modules we bought off Ebay were fakes: they were not
made by Jnhuamao \footnote{Jnhuamao is the company which makes the HM-10
  module, along with other Bluetooth modules. Their English website is:
  \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.}, and did not come
with genuine Jnhuamao firmware. However, they did have real TI CC2541
chips. We realized they were fakes when they did not behave according to
the Jnhuamao data sheet (see \protect\hyperlink{data-sheets}{the data
sheets section}). Luckily, the hardware on the fake chips is the same as
that of the genuine chips, minus an external crystal, and the genuine
firmware checks for the presence of the crystal, and works even without
it. \footnote{\url{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}.}
As such, we reprogrammed the chips with the genuine firmware according
to an
\href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
forum post}:

@@

After flashing genuine firmware onto the chip, the next step was to
update the firmware to the latest version. The firmware flashed onto the
board was version 540, but Jnhuamao had (at the time we did this
project)
\href{http://www.jnhuamao.cn/rom/HMSoft-10-2541-V603.zip}{released
version 603}. \footnote{Jnhuamao's firmware download page
  \url{http://www.jnhuamao.cn/download_rom_en.asp?id=1\#}.} They also
provide
\href{http://www.jnhuamao.cn/HowToUpgradeFirmware_en.zip}{instructions
on how to upgrade the firmware}. The basic process is:

@@

\begin{verbatim}
OK+DISC:00000000:00000000000000000000000000000000:0000000000:6832A3801EBE:-080
\end{verbatim}

@@

The PIC commmunicates with the IMU via I2C. The IMU includes a breakout
board for the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module, which itself includes 2 dies. One
contains the 3-axis gyroscope and 3-axis accelerometer, which were not
used in this project, and the other die is the AK8963 3-axis
magnetometer (compass).

@@

@@

where \(A\) and \(n\) are RF propagation parameters in dBm, \(d\) is
distance in meters, and RSSI is the measured RSSI in dBm. \footnote{L.
  Peneda, A. Azenha and A. Carvalho, ``Trilateration for indoors
  positioning within the framework of wireless communications,'' 2009
  35th Annual Conference of IEEE Industrial Electronics, Porto, 2009,
  pp.~2732-2737.

  doi: 10.1109/IECON.2009.5415423

  URL:
  \url{http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=5415423\&isnumber=5414636}}
We experimented with RSSI measurements to determine how well they worked
by taking 2 Bluetooth modules, and measuring the RSSI while changing the
distance between them. One remained stationary on the floor and the
other was moved away from it 1 floor tile (each floor tile is a 1 foot
square) at a time. At each point, we took 3 RSSI measurements and
averaged them. The graph below displays the results:

@@

We reused some of the code from the
\href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/index.html}{Self-Balancing
Robot} as
\protect\hyperlink{code-and-designs-borrowed-from-others}{cited} in our
references section, and explained in \protect\hyperlink{imu}{IMU}. The
rest of the design is ours, and is not an infringement on intellectual
property. We also used the Bluetooth firmware flashing technique we
discovered on the
\href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
forums}.

We did not reverse-engineer any technology, and did not encounter any
issues with trademarks or patents. We did not have to sign
non-disclosure to get a sample part. We hope to further polish our
findings and report, and will aim to publish them in a print magazine,
ideally as a part of ECE 4920.

@@

\hypertarget{references}{%
\subsubsection{References}\label{references}}

\hypertarget{data-sheets}{%
\paragraph{Data sheets}\label{data-sheets}}

\begin{itemize}
\tightlist
\item
  \href{https://www.akm.com/akm/en/file/datasheet/AK8963C.pdf}{3-axis
  Electronic Compass on IMU}.
\item
  \href{http://www.jnhuamao.cn/bluetooth40_en.zip}{HM-10 (Bluetooth
  breakout module \& firmware)}. The ZIP contains the original data
  sheet, as well as other documentation.
  \href{http://fab.cba.mit.edu/classes/863.15/doc/tutorials/programming/bluetooth/bluetooth40_en.pdf}{MIT
  has a PDF} available, which may me out of date, but is easier to get
  to.
\item
  \href{http://ww1.microchip.com/downloads/en/DeviceDoc/60001168J.pdf}{PIC32MX250}.
\end{itemize}

\hypertarget{code-and-designs-borrowed-from-others}{%
\paragraph{Code and designs borrowed from
others}\label{code-and-designs-borrowed-from-others}}

\begin{itemize}
\tightlist
\item
  \href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/index.html}{Self-Balancing
  Robot} by Desmond Caulley (dc686@cornell.edu), Nadav Nehoran
  (nn233@cornell.edu), Sherry Zhao (hz263@cornell.edu). In particular,
  we borrowed extensively from their
  \href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/dc686_nn233_hz263/i2c_helper.h}{i2c\_helper.h}.
\end{itemize}

\hypertarget{other}{%
\paragraph{Other}\label{other}}

\begin{itemize}
\tightlist
\item
  \href{http://people.ece.cornell.edu/land/courses/ece4760/}{ECE 4760
  Website} by Bruce Land
\item
  Website template:
  \url{https://github.com/tajmone/pandoc-goodies/tree/master/templates/html5/github}.
\end{itemize}

>>>>

The HM-10 bluetooth modules we bought off Ebay were fakes: they were not
made by Jnhuamao \footnote{Jnhuamao is the company which makes the HM-10
  module, along with other Bluetooth modules. Their English website is:
  \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.}, and did not come
with genuine Jnhuamao firmware. However, they did have real TI CC2541
chips. We realized they were fakes when they did not behave according to
the Jnhuamao data sheet. \footnote{The original data sheet comes from \cite{jnhuamaodatasheet}. A PDF is available at \cite{jnhuamaomit}, which may be out of date, but is eaiser to get.}
Luckily, the hardware on the fake chips is the same as
that of the genuine chips, minus an external crystal, and the genuine
firmware checks for the presence of the crystal, and works even without
it. \cite{crystal}
As such, we reprogrammed the chips with the genuine firmware according
to an
\href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
forum post}  \cite{crystal}:

@@

After flashing genuine firmware onto the chip, the next step was to
update the firmware to the latest version. The firmware flashed onto the
board was version 540, but Jnhuamao had (at the time we did this
project) released
version 603. \cite{jnhuamao603} Based on their instructions \cite{jnhuamaoinstructions}, we flashed our chips:

@@

\begin{verbatim}
OK+DISC:00000000:00000000000000000000000000000000:
  0000000000:6832A3801EBE:-080
\end{verbatim}

@@

The PIC commmunicates with the IMU via I2C. The IMU includes a breakout
board for the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module, which itself includes 2 dies. One
contains the 3-axis gyroscope and 3-axis accelerometer, which were not
used in this project, and the other die is the AK8963 3-axis
magnetometer (compass). \cite{ak8963cdatasheet}

@@

We based our IMU code off of the \texttt{i2c\_helper.h} from the Self-Balancing Robot project. \cite{selfbalancingrobot}

@@

where \(A\) and \(n\) are RF propagation parameters in dBm, \(d\) is
distance in meters, and RSSI is the measured RSSI in dBm. \cite{5415423}
We experimented with RSSI measurements to determine how well they worked
by taking 2 Bluetooth modules, and measuring the RSSI while changing the
distance between them. One remained stationary on the floor and the
other was moved away from it 1 floor tile (each floor tile is a 1 foot
square) at a time. At each point, we took 3 RSSI measurements and
averaged them. The graph below displays the results:

@@

@@
